---
title: "Untitled"
author: "William Lai"
date: "11 May 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(stringr)
library(quanteda)
```

```{r "Parameter"}
gamma2 <- 0.5  # bigram discount
gamma3 <- 0.5  # trigram discount
#bigPre <- 'happi_new'
# Q1
#q <- "The guy in front of me just bought a pound of bacon, a bouquet, and a case of"
#filterTxt <- c("soda", "pretzels", "beer", "cheese")
# Q2
#q <- "The guy in front of me just bought a pound of bacon, a bouquet, and a case of"
#filterTxt <- c("world", "universe", "best", "most")
# Q3
#q <- "Hey sunshine, can you follow me and make me the"
#filterTxt <- c("happiest", "saddest", "smelliest", "bluest")
# Q4
q <- "Very early observations on the Bills game: Offense still struggling but the"
filterTxt <- c("players", "referees", "defense", "crowd")
# Q5
q <- "Go on a romantic date at the"
filterTxt <- c("grocery", "movies", "mall", "beach")
# Q6
q <- "Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my"
filterTxt <- c("way", "motorcycle", "horse", "phone")
# Q7
q <- "Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some"
filterTxt <- c("years", "thing", "time", "weeks")
# Q8
q <- "After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little"
filterTxt <- c("ears", "fingers", "toes", "eyes")
# Q9
q <- "Be grateful for the good times and keep the faith during the"
filterTxt <- c("worse", "hard", "bad", "sad")
# Q10
q <- "If this isn't the cutest thing you've ever seen, then you must be"
filterTxt <- c("callous", "asleep", "insane", "insensitive")

toks <- tokens(q, to_lower = TRUE, remove_punct = TRUE, remove_symbols = TRUE, remove_twitter = TRUE, remove_numbers = TRUE)
toks <- tokens_remove(toks, stopwords('en'))
tok <- unlist(toks)
bigPre <- paste(tok[length(tok) - 1], tok[length(tok)], sep = "_")
```

```{r "n-grams", cache = TRUE}
load("freq.rds")
unigs <- uni
bigrs <- bi
trigs <- tri
```

## Step 3. Calculate Probabilities of Words Completing Observed Trigrams

```{r "3"}
## Returns a two column data.frame of observed trigrams that start with the
## bigram prefix (bigPre) in the first column named ngram and
## frequencies/counts in the second column named freq. If no observed trigrams
## that start with bigPre exist, an empty data.frame is returned.
##
## bigPre -  single-element char array of the form w2_w1 which are the first 
##           two words of the trigram we are predicting the tail word of
## trigrams - 2 column data.frame or data.table. The first column: ngram,
##            contains all the trigrams in the corpus. The second column:
##            freq, contains the frequency/count of each trigram.
getObsTrigs <- function(bigPre, trigrams) {
  trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
                           freq=vector(mode = 'integer', length = 0))
  regex <- sprintf("%s%s%s", "^", bigPre, "_")
  trigram_indices <- grep(regex, trigrams$ngram)
  if(length(trigram_indices) > 0) {
    trigs.winA <- trigrams[trigram_indices, ]
  }
  
  return(trigs.winA)
}

## Returns a two column data.frame of observed trigrams that start with bigram
## prefix bigPre in the first column named ngram and the probabilities
## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
##
## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
##            contains all the observed trigrams that start with the bigram
##            prefix bigPre which we are attempting to the predict the next
##            word of in a give phrase. The second column: freq, contains the
##            frequency/count of each trigram.
## bigrs - 2 column data.frame or data.table. The first column: ngram,
##         contains all the bigrams in the corpus. The second column:
##         freq, contains the frequency/count of each bigram.
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## triDisc - amount to discount observed trigrams
getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
  if(nrow(obsTrigs) < 1) return(NULL)
  obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
  obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
  colnames(obsTrigProbs) <- c("ngram", "prob")
  
  return(obsTrigProbs)
}

obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
# convert counts to probabilities
qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
qbo_obs_trigrams
```

## Step 4. Calculate Probabilities of Words Completing Unobserved Trigrams

### step 4. i. Find unobserved trigram tail words:

```{r "4.1"}
## Returns a character vector which are the tail words of unobserved trigrams
## that start with the first two words of obsTrigs (aka the bigram prefix).
## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
## describing the details of equation 17.
##
## obsTrigs - character vector of observed trigrams delimited by _ of the form:
##            w3_w2_w1 where w3_w2 is the bigram prefix
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
getUnobsTrigTails <- function(obsTrigs, unigs) {
  obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
  unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
  return(unobs_trig_tails)
}

unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
#unobs_trig_tails
head(unobs_trig_tails, 10)
```

### Step 4. ii. Calculate discounted probability mass at the bigram level α(wi−1):

```{r "4.2"}
## Returns the total probability mass discounted from all observed bigrams
## calculated from equation 14.  This is the amount of probability mass which
## is redistributed to UNOBSERVED bigrams. If no bigrams starting with
## unigram$ngram[1] exist, 0 is returned.
##
## unigram - single row, 2 column frequency table. The first column: ngram,
##           contains the w_i-1 unigram (2nd word of the bigram prefix). The
##           second column: freq, contains the frequency/count of this unigram.
## bigrams - 2 column data.frame or data.table. The first column: ngram,
##           contains all the bigrams in the corpus. The second column:
##           freq, contains the frequency or count of each bigram.
## bigDisc - amount to discount observed bigrams
getAlphaBigram <- function(unigram, bigrams, bigDisc=0.5) {
  # get all bigrams that start with unigram
  regex <- sprintf("%s%s%s", "^", unigram$ngram[1], "_")
  bigsThatStartWithUnig <- bigrams[grep(regex, bigrams$ngram),]
  if(nrow(bigsThatStartWithUnig) < 1) return(0)
  alphaBi <- 1 - (sum(bigsThatStartWithUnig$freq - bigDisc) / unigram$freq)
  
  return(alphaBi)
}

unig <- str_split(bigPre, "_")[[1]][2]
unig <- unigs[unigs$ngram == unig,]
alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
alpha_big
```

### Step 4. iii. Calculate backed off probabilities qBO for bigrams

```{r "4.3.1"}
## Returns a character vector of backed off bigrams of the form w2_w1. These 
## are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
## prefix bigPre and w are the tail words of unobserved bigrams that start with
## w_i-1.
##
## bigPre - single-element char array of the form w2_w1 which are first two
##          words of the trigram we are predicting the tail word of
## unobsTrigTails - character vector that are tail words of unobserved trigrams
getBoBigrams <- function(bigPre, unobsTrigTails) {
  w_i_minus1 <- str_split(bigPre, "_")[[1]][2]
  boBigrams <- paste(w_i_minus1, unobsTrigTails, sep = "_")
  return(boBigrams)
}

## Returns a two column data.frame of backed-off bigrams in the first column
## named ngram and their frequency/counts in the second column named freq.
## 
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## unobsTrigTails - character vector that are tail words of unobserved trigrams
## bigrs - 2 column data.frame or data.table. The first column: ngram,
##         contains all the bigrams in the corpus. The second column:
##         freq, contains the frequency/count of each bigram.
getObsBoBigrams <- function(bigPre, unobsTrigTails, bigrs) {
  boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
  obs_bo_bigrams <- bigrs[bigrs$ngram %in% boBigrams, ]
  return(obs_bo_bigrams)
}

## Returns a character vector of backed-off bigrams which are unobserved.
##
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## unobsTrigTails - character vector that are tail words of unobserved trigrams
## obsBoBigram - data.frame which contains the observed bigrams in a column
##               named ngram
getUnobsBoBigrams <- function(bigPre, unobsTrigTails, obsBoBigram) {
  boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
  unobs_bigs <- boBigrams[!(boBigrams %in% obsBoBigram$ngram)]
  return(unobs_bigs)
}

## Returns a dataframe of 2 columns: ngram and probs.  Values in the ngram
## column are bigrams of the form: w2_w1 which are observed as the last
## two words in unobserved trigrams.  The values in the prob column are
## q_bo(w1 | w2) calculated from from equation 10.
##
## obsBoBigrams - a dataframe with 2 columns: ngram and freq. The ngram column
##                contains bigrams of the form w1_w2 which are observed bigrams
##                that are the last 2 words of unobserved trigrams (i.e. "backed
##                off" bigrams). The freq column contains integers that are
##                the counts of these observed bigrams in the corpus.
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
## bigDisc - amount to discount observed bigrams
getObsBigProbs <- function(obsBoBigrams, unigs, bigDisc=0.5) {
  first_words <- str_split_fixed(obsBoBigrams$ngram, "_", 2)[, 1]
  first_word_freqs <- unigs[unigs$ngram %in% first_words, ]
  obsBigProbs <- (obsBoBigrams$freq - bigDisc) / first_word_freqs$freq
  obsBigProbs <- data.frame(ngram=obsBoBigrams$ngram, prob=obsBigProbs)
  
  return(obsBigProbs)
}

## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
## column are unobserved bigrams of the form: w2_w1.  The values in the prob
## column are the backed off probability estimates q_bo(w1 | w2) calculated
## from from equation 16.
##
## unobsBoBigrams - character vector of unobserved backed off bigrams
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
## alphaBig - total discounted probability mass at the bigram level
getQboUnobsBigrams <- function(unobsBoBigrams, unigs, alphaBig) {
  # get the unobserved bigram tails
  qboUnobsBigs <- str_split_fixed(unobsBoBigrams, "_", 2)[, 2]
  w_in_Aw_iminus1 <- unigs[!(unigs$ngram %in% qboUnobsBigs), ]
  # convert to data.frame with counts
  qboUnobsBigs <- unigs[unigs$ngram %in% qboUnobsBigs, ]
  denom <- sum(qboUnobsBigs$freq)
  # converts counts to probabilities
  qboUnobsBigs <- data.frame(ngram=unobsBoBigrams,
                             prob=(alphaBig * qboUnobsBigs$freq / denom))
  
  return(qboUnobsBigs)
}

bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)  # get backed off bigrams
# separate bigrams which use eqn 10 and those that use 16
obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
# unobs_bo_bigrams = c("the_buy", "the_EOS", "the_paint", "the_sell", "the_the")
# calc obs'd bigram prob's from eqn 10
qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2) #ngram     probs
# calc alpha_big & unobs'd bigram prob's from eqn 16             #the_house 0.3125
unig <- str_split(bigPre, "_")[[1]][2]
unig <- unigs[unigs$ngram == unig,]
# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
qbo_bigrams <- rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
#qbo_bigrams
head(qbo_bigrams, 10)
```

#### Checking the Bigram Calculations

```{r "4.3.2"}
unobs <- qbo_bigrams[-1,]
sum(unobs$prob)
```

### Step 4. iv. Calculate discounted probability mass at the trigram level α(wi−2,wi−1)

```{r "4.4"}
## Returns the total probability mass discounted from all observed trigrams.
## calculated from equation 14. This is the amount of probability mass which is
## redistributed to UNOBSERVED trigrams. If no trigrams starting with
## bigram$ngram[1] exist, 1 is returned.
##
## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
##            contains all the observed trigrams that start with the bigram
##            prefix we are attempting to the predict the next word of. The 
##            second column: freq, contains the frequency/count of each trigram.
## bigram - single row frequency table where the first col: ngram, is the bigram
##          which are the first two words of unobserved trigrams we want to
##          estimate probabilities of (same as bigPre in other functions listed
##          prior) delimited with '_'. The second column: freq, is the
##          frequency/count of the bigram listed in the ngram column.
## triDisc - amount to discount observed trigrams
getAlphaTrigram <- function(obsTrigs, bigram, triDisc=0.5) {
  if(nrow(obsTrigs) < 1) return(1)
  alphaTri <- 1 - sum((obsTrigs$freq - triDisc) / bigram$freq[1])
  
  return(alphaTri)
}

bigram <- bigrs[bigrs$ngram %in% bigPre, ]
alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
#alpha_trig
head(alpha_trig, 10)
```

### Step 4. v. Calculate unobserved trigram probabilities qBO(wi|wi−2,wi−1)

```{r "4.5"}
## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
## column are unobserved trigrams of the form: w3_w2_w1.  The values in the prob
## column are q_bo(w1 | w3, w2) calculated from equation 17.
##
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## qboObsBigrams - 2 column data.frame with the following columns -
##                 ngram: observed bigrams of the form w2_w1
##                 probs: the probability estimate for observed bigrams:
##                        qbo(w1 | w2) calc'd from equation 10.
## qboUnobsBigrams - 2 column data.frame with the following columns -
##                   ngram: unobserved bigrams of the form w2_w1
##                   probs: the probability estimate for unobserved bigrams
##                          qbo(w1 | w2) calc'd from equation 16.
## alphaTrig - total discounted probability mass at the trigram level
getUnobsTriProbs <- function(bigPre, qboObsBigrams,
                             qboUnobsBigrams, alphaTrig) {
  qboBigrams <- rbind(qboObsBigrams, qboUnobsBigrams)
  qboBigrams <- qboBigrams[order(-qboBigrams$prob), ]
  sumQboBigs <- sum(qboBigrams$prob)
  first_bigPre_word <- str_split(bigPre, "_")[[1]][1]
  unobsTrigNgrams <- paste(first_bigPre_word, qboBigrams$ngram, sep="_")
  unobsTrigProbs <- alphaTrig * qboBigrams$prob / sumQboBigs
  unobsTrigDf <- data.frame(ngram=unobsTrigNgrams, prob=unobsTrigProbs)
  
  return(unobsTrigDf)
}

qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
                                       qbo_unobs_bigrams, alpha_trig)
#qbo_unobs_trigrams
head(qbo_unobs_trigrams, 10)

unobs <- qbo_unobs_trigrams %>% transform(first_term = sub("(.*)_(.*)", "\\1", ngram), second_term = sub("(.*)_(.*)", "\\2", ngram)) %>% select(first_term, second_term, prob) 
#save(unobs, file = "unobs.rds")
```

## Step 5. Select wi with the highest qBO(wi|wi−2,wi−1)

```{r "5"}
getPredictionMsg <- function(qbo_trigs) {
    # pull off tail word of highest prob trigram
    prediction <- str_split(qbo_trigs$ngram[1], "_")[[1]][3]
    result <- sprintf("%s%s%s%.4f", "highest prob prediction is >>> ", prediction,
                      " <<< which has probability = ", qbo_trigs$prob[1])
    return(result)
}

qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]  # sort by desc prob
out_msg <- getPredictionMsg(qbo_trigrams)
#out_msg
head(out_msg, 10)
```

```{r}
q
bigPre
unobs %>% filter(second_term %in% filterTxt) %>% arrange(desc(prob))
```